---
title: "flowDashboard Vignette"
author: "Ted Laderas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document outlines how to set up a flow cytometry dashboard using the `flowDashboard` Shiny modules.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Where `flowDashboard` Fits in an Analysis Workflow

`flowDashboard` enables you to visualize data at the Sample and Population Level. The package assumes that you either have FCS data in a directory, or have generated a GatingSet for your gating scheme. This gating scheme can either be automated (from a pipeline such as `openCyto`) or manually (imported from flowJo). 

There are multiple modules that will enable you to examine each step of the analysis. 

  - `Annotation` display - displays annotation for all the FCS files in an experiment.
  - `QC` plots - plots that let you compare the overall state of the data before and after preprocessing.
  - `Gating` plots - plots that display how the data was gated
  - `Dot` plots - to visualize populations based on a covariate
  - `Violin` plots of markers in a population
  - `Waterfall` plots to examine how population percentage correlates with an outcome.
  - `Coexpression` plots - a heatmap and a two-way expression explorer.

## Preprocessing Data

Once your data is in a GatingSet (preferably in ncdf format), we will need to transform and prep it into the `data.table` format. This package uses `data.table` because it is a very fast format for accessing and displaying data.

The annotation structure is another `data.table` that we need for the dashboard. Based on your unit of analysis and what the FCS file represents (which may be patients, cell lines, or conditions within a patient), you will need an annotation table that provides covariate information at the level of analysis you are interested in. You can generate the skeleton of an annotation table with the .

## QC Module

The QC Heatmap provides a quick overview of how markers vary across populations. This heatmap is based on the idea of examining a QC panel of markers to compare whether there is a large technical variation between samples.

The QC Violin plots allow you to examine 

The underlying QC data is sampled using the `buildSampledDataList()` function. This function generates a data.table that samples from each FCS file for a selected control panel of markers.

## Gating Module

The gating module consists of a heatmap visualization and plots for each cell in the heatmap, corresponding to the gating scheme. This module is useful in understanding the impact of automated gating pipelines, such as `openCyto`, and comparing manual gating schemes across samples. 

The plots for each sample must be precomputed using the `plotAllPopulations()` function. This function takes a GatingSet as input.

The Violin plots 

## Coexpression Module

## Waterfall Module

## DotPlot Module


## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
