---
title: "flowDashboard Vignette"
author: "Ted Laderas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document outlines how to set up a flow cytometry dashboard using the `flowDashboard` Shiny modules.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "../")
```

## Where `flowDashboard` Fits in an Analysis Workflow

`flowDashboard` enables you to visualize data at both the Sample and cell population Level. The package assumes that you either have FCS data in a directory, or have generated a `GatingSet` for your gating scheme. This gating scheme can either be automated (from a pipeline such as `openCyto`) or manually (imported from flowJo). 

There are multiple modules that will enable you to examine each step of the analysis. 

  - `Annotation` display - displays annotation for all the FCS files in an experiment.
  - `QC` plots - plots that let you compare the overall state of the data before and after preprocessing.
  - `Gating` plots - plots that display how the data was gated
  - `Dot` plots - to visualize populations based on a covariate
  - `Violin` plots of markers in a population
  - `Waterfall` plots to examine how population percentage correlates with an outcome.
  - `Coexpression` plots - a heatmap and a two-way expression explorer.

## Preprocessing Data

Once your data is in a `gatingSet` (preferably in ncdf format), we will need to transform and prep it into the `data.table` format. This package uses `data.table` because it is a very fast format for accessing, merging, sorting, and displaying data.

The annotation structure is another `data.table` that we need for the dashboard. Based on your unit of analysis and what the FCS file represents (which may be patients, cell lines, or conditions within a patient), you will need an annotation table that provides covariate information at the level of analysis you are interested in. You can generate the skeleton of an annotation table with the `` command, or you can use the phenoData slot in your GatingSet.

Please note that the usefulness of `flowDashboard` is only as good as your annotation. Careful curation of variables in your annotation will allow you to sort, subset, and aggregate on these variables.

```{r}
library(flowWorkspace)
library(flowDashboard)
gsFile <- system.file("extdata/gvHDgs/", package = "flowDashboard")
gs <- load_gs(gsFile)

##look at phenoData slot in gs
phenotypes <- pData(gs@data@phenoData)
phenotypes

##look at marker information in gs
markerInfo <- pData(parameters(gs@data[[1]]))
markerInfo

#build a qcFlowObj 
QCO <- QCOFromGatingSet(gs, samplePop = 1000)

#show annotation slot
QCO$annotation

#show first few lines of qcData
QCO$qcData
```


## QC Module

The QC Heatmap provides a quick overview of how markers vary across populations. This heatmap is based on the idea of examining a QC panel of markers to compare whether there is a large technical variation between samples.

The QC Violin plots allow you to compare fluoresence distibution of individual functions.

The underlying QC data is sampled using the `buildSampledDataList()` function. This function generates a data.table that samples from each FCS file for a selected control panel of markers.

## Gating Module

The gating module consists of a heatmap visualization and plots for each cell in the heatmap, corresponding to the gating scheme. This module is useful in understanding the impact of automated gating pipelines, such as `openCyto`, and comparing manual gating schemes across samples. 

The plots for each sample must be precomputed using the `plotAllPopulations()` function. This function takes a GatingSet as input.

The Violin plots 

## Coexpression Module

## Waterfall Module

This module displays population percentages 

## DotPlot Module


